---
title: "RcppParallel vector of std::string"
author: "Brian J. Knaus"
date: "March 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


I work with tabular or matrix-like data where samples are in columns and traits (variants) are in rows.
However, each value of this matrix, or record, contains multiple pieces of information that is delimited.
An example is as follows.


$$
0|0:48:1:51,51
$$


Note that while the data technically is numeric, because it is delimited we need to handle it in a non-numeric container, such as `std::string`.
Also note that while the data is delimited with a colon (`:`) there is secondary delimitation.
For example, the fourth element is comma delimited.
This means it would be nice to engineer a flexible solution.
Because each row is an independent trait, it seemed like a good candidate for parallelization.
Because I already use `Rcpp` it seemed that `RcppParallel` would be a good solution.
The `R` and `Rcpp` data structures are not thread safe because a garbage collection event may occur.
To solve this, `RcppParallel` provides `RVector` and `RMatrix`.
However, these are only suited for numeric data.
A post to `rcpp-devel` suggested I could use C++ containers as a thread safe alternative.
Ultimately I would like to work on some form of matrix (e.g., std::vector< std::vector< std::string > >) but here I work with a `std::vector< std::string > >` in order to simplify the task.
Here I attempt to implement multi-threading using `RcppParallel` and `std::vector< std::string >`.


## Implementation using Rcpp


```{r Rcpp, engine='Rcpp'}
#include <Rcpp.h>


std::string getElement( std::string mystring, char split, int element ) {
  std::string rstring;
  int elementN = 0;
  unsigned int i=0;
  int start = 0;
  
  for(i=1; i<mystring.size(); i++){
    if(mystring[i] == split){
      // We matched our splitting character.
      if(elementN == element){
        // We've arrived at our desired element.
        rstring = mystring.substr(start, i - start);
        return rstring;
      } else {
        // We need to iterate to the next element.
        start = i + 1;
        i = i + 1;
        elementN = elementN + 1;
      }
    }
  }
  
  // The end of the string has no delimiter.
  // We test if it is the correct element and parse if so.
  if(elementN == element){
    rstring = mystring.substr(start, i - start);
    return rstring;
  } else {
    rstring = "-999";
    return rstring;
  }
}

// [[Rcpp::export]]
Rcpp::StringVector delimitString( Rcpp::StringVector myvector, char myDelim, int element ) {
  Rcpp::StringVector rvector( myvector.size() );
  std::string myString;
  
  // R is one based, C++ is zero based.
  element = element - 1;
  
  unsigned int i=0;
  for(i=0; i<myvector.size(); i++){
    myString = myvector[i];
    rvector[i] = getElement(myString, myDelim, element);
  }
  
  return rvector;
}


```



## Parallel version using RcppParallel

```{r RcppParallel}
#
Rcpp::sourceCpp('rcppparallel_string.cpp')
```


## Benchmarking


First we'll fabricate some example data.


```{r data}
#n <- 1e7
#n <- 1e6
#
n <- 1e1
myGT <- paste(sample(0:1, size = n, replace = TRUE),
              sample(0:1, size = n, replace = TRUE),
              sep = "/")
myGT <- paste(myGT, LETTERS, 
              paste(sample(1:200, size = n, replace = TRUE), 
                    sample(1:200, size = n, replace = TRUE), sep=","),
              sep = ":")
myGT <- paste(myGT, sample(1:200, size = n, replace = TRUE), sep = ":")
head(myGT)
```





```{r implement Rcpp}
library(microbenchmark)
nTimes <- 20
microbenchmark(gt1 <- delimitString(myGT, ":", 3), times = nTimes)
head(gt1)
```


```{r implement RcppParallel}
#
microbenchmark(gt2 <- rcpp_parallel_delimitString(myGT), times = nTimes)
#
head(gt2)
```


